Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2018-10-02 paper/paper.tex
made legend to Figure 7 more succinct

\figsupp[Validation of viral titer normalization for \FIG{validation}B.]
{Validation that the infections in \FIG{validation}B and \FIGSUPP[validation]{delflow} were performed at similar doses of virions capable of initiating primary transcription.
In this experiment, A549 cells were infected at MOI of 0.4 (based on TCID50 as described in \FIGSUPP[validation]{delflow}), and then after 8 hours mRNA was harvested for qPCR on oligo-dT primed reverse transcription products.
The y-axis shows the ratio of viral HA mRNA to the housekeeping gene L32.
These infections were performed in the presence of absence of 50 $\mu$g/ml cycloheximide, which blocks protein synthesis and hence secondary transcription by newly synthesized viral proteins.
In the absence of cycloheximide, the viruses with deletions in PB1 produced less viral mRNA presumably because they could not produce PB1 protein for secondary transcription.
But in the presence of cycloheximide, all viruses produced similar amounts of viral mRNA, indicating that the dose of particles active for primary transcription is roughly equivalent across variants.
Each measurement was performed in quadruplicate.
}
{\includegraphics[width=0.6\textwidth]{figures/Validation_Figure/qpcr_plot.pdf}}
\label{figsupp:delqPCR}

@@

\centerline{\includegraphics[width=0.85\textwidth]{figures/MutantComparison/p_merged.pdf}}
\caption{
For NS1-deficient infections, higher viral gene expression causes more IFN induction.
{\bf (A)}
Flow cytometry shows that for viruses with defects in NS1, there is increased IFN induction among infected cells that express high levels of HA protein relative to those that express low levels of HA.
However, for wildtype virus and viruses with mutations in PB1, there is only a modest increase in IFN induction with increasing HA.
The y-axis shows the ratio of the percent of cells that are IFN+ in the highest HA-expression quartile relative to the lowest HA-expression quartile.
Points indicate four replicates, and lines indicate the mean.
See \FIGSUPP[mutantcomparison]{ha_vs_ifn} for details.
{\bf (B)}
Higher viral gene expression \emph{causes} more IFN induction for NS1-deficient infections, since blocking viral secondary transcription strongly decreases IFN levels in cells infected with NS1stop virus, but not cells infected with the virus carrying a deletion in PB1.
A549 cells were infected with both viral stocks at doses that induced equivalent IFN, and ribavirin was added to half the samples at a concentration (100 $\mu$M) sufficient to block secondary transcription~\citep{Vanderlinden:2016ec,Reuther:2015ef,Scholtissek:1976wg}.
After eight hours, IFNB1 transcript was quantified relative to the housekeeping gene L32 by qPCR. 
Points indicate four replicates, and lines indicate the mean.
The dashed gray line indicates the limit of detection.
\FIGSUPP[mutantcomparison]{cycloheximide} shows that similar results are obtained if secondary transcription is blocked with cycloheximide rather than ribavirin.
}
\label{fig:mutantcomparison}

\figsupp[Flow cytometry analysis showing that viral burden strongly correlates with IFN induction in NS1-deficient infections.]
{Re-analysis of the flow cytometry data in \FIGSUPP[validation]{SNPflow} indicates that a higher viral burden is associated with increased IFN-induction in NS1-deficient infections.
We re-analyzed all the infected cells (defining ``infected'' using the same HA-gating scheme in \FIGSUPP[validation]{SNPflow}) for the wildtype virus, the NS1stop virus, and the three amino-acid mutants with significantly increased IFN induction.
For each virus and replicate, we binned the infected cells into HA expression quartiles.
We then calculated the percent of cells that were IFN+ in each quartile (defining ``IFN+'' using the same gating scheme in \FIGSUPP[validation]{SNPflow}).
The plots show the mean HA expression in each quartile versus the percent of cells that are IFN+.
The results clearly indicate that for the NS1stop virus, increased viral burden (more HA) correlates with IFN induction.
A similar trend also holds for the virus with an amino-acid substitution in NS1 (NS1-A122V). 
But for wildtype and the two PB1 mutants, there is no clear trend for higher viral burden to correlate with IFN induction.
\FIG{mutantcomparison}A summarizes the data in this figure by simply showing the ratio of the percent IFN+ of the highest HA-expression quartile to the lowest HA-expression quartile for each sample and replicate.}
{\includegraphics[width=0.9\textwidth]{figures/MutantComparison/p_ifn_vs_ha.pdf}}
\label{figsupp:ha_vs_ifn}

\figsupp[Similar to \FIG{mutantcomparison}B, but uses cycloheximide rather than ribavirin.]
{Preventing viral secondary transcription by addition of the cycloheximide also decreases the rate of IFN induction by the NS1stop virus relative to the PB1 deletion virus, corroborating the result in \FIG{mutantcomparison}B with a different inhibitor.
However, the results are more complex here, since cycloheximide alone increases IFN production~\citep{raj1981analysis, maroteaux1983cycloheximide, killip2014activation}.
Therefore, we see an increase in IFN levels upon cycloheximide treatment for the uninfected cells and cells infected with the PB1 deletion virus, but a decrease in cells infected with the NS1stop virus.
This plot shows qPCR results for the same samples as in \FIGSUPP[validation]{cycloheximide}, but with the qPCR for expression of IFNB1 transcripts relative to the housekeeping gene L32.
As described in \FIGSUPP[validation]{cycloheximide}, A549 cells were infected at an MOI of 0.4, such that HA expression was equivalent between viral variants when secondary transcription was blocked, and some of the samples treated with 50 $\mu$g/ml of cycloheximide.
After 8 hours, transcript levels were quantified by qPCR.
Points indicate four replicates, and lines indicate the mean.
The dashed line shows the limit of detection, and points below this limit where set to that value.
}
{\includegraphics[width=0.6\textwidth]{figures/MutantComparison/cyclo_qPCR_plot.pdf}}
\label{figsupp:cycloheximide}

>>>>

\figsupp[Validation of viral titer normalization for \FIG{validation}B.]
{Validation that the infections in \FIG{validation}B and \FIGSUPP[validation]{delflow} were performed at similar doses of virions capable of initiating primary transcription.
In this experiment, A549 cells were infected at MOI of 0.4 (based on TCID50 as described in \FIGSUPP[validation]{delflow}), and then after 8 hours mRNA was harvested for qPCR on oligo-dT primed reverse transcription products.
The y-axis shows the ratio of viral HA mRNA to the housekeeping gene L32.
These infections were performed in the presence of absence of 50 $\mu$g/ml cycloheximide, which blocks protein synthesis and hence secondary transcription by newly synthesized viral proteins~\citep{killip2014activation}.
In the absence of cycloheximide, the viruses with deletions in PB1 produced less viral mRNA presumably because they could not produce PB1 protein for secondary transcription.
But in the presence of cycloheximide, all viruses produced similar amounts of viral mRNA, indicating that the dose of particles active for primary transcription is roughly equivalent across variants.
Each measurement was performed in quadruplicate.
}
{\includegraphics[width=0.6\textwidth]{figures/Validation_Figure/qpcr_plot.pdf}}
\label{figsupp:delqPCR}

@@

\centerline{\includegraphics[width=0.85\textwidth]{figures/MutantComparison/p_merged.pdf}}
\caption{
For NS1-deficient infections, higher viral gene expression causes more IFN induction.
{\bf (A)}
Flow cytometry shows that only for viruses with defects in NS1, there is more IFN induction among infected cells that express higher levels of HA protein.
The y-axis shows the ratio of the percent of IFN+ cells in the highest HA-expression quartile relative to the lowest HA-expression quartile.
Points indicate replicates, and lines indicate the mean.
See \FIGSUPP[mutantcomparison]{ha_vs_ifn} for details.
{\bf (B)}
Blocking viral secondary transcription strongly decreases IFN levels in cells infected with NS1stop virus, but not cells infected with virus carrying a deletion in PB1.
A549 cells were infected with the viral stocks at doses that induced equivalent IFN, and ribavirin was added at a concentration (100 $\mu$M) sufficient to block secondary transcription~\citep{Vanderlinden:2016ec,Reuther:2015ef,Scholtissek:1976wg}.
After eight hours, IFNB1 transcript was quantified relative to the housekeeping gene L32 by qPCR. 
The dashed gray line indicates the limit of detection.
\FIGSUPP[mutantcomparison]{cycloheximide} shows that results are similar if secondary transcription is instead blocked with cycloheximide.
}
\label{fig:mutantcomparison}

\figsupp[More detailed data on the flow cytometry analysis in \FIG{mutantcomparison}A.]
{Analysis of the flow cytometry data in \FIGSUPP[validation]{SNPflow} indicates that a higher viral burden is associated with increased IFN-induction in NS1-deficient infections.
We re-analyzed all the infected cells (defining ``infected'' using the same HA-gating scheme in \FIGSUPP[validation]{SNPflow}) for the wildtype virus, the NS1stop virus, and the three amino-acid mutants with significantly increased IFN induction.
For each virus and replicate, we binned the infected cells into HA expression quartiles.
We then calculated the percent of cells that were IFN+ in each quartile (defining ``IFN+'' using the same gating scheme in \FIGSUPP[validation]{SNPflow}).
The plots show the mean HA expression in each quartile versus the percent of cells that are IFN+.
The results clearly indicate that for the NS1stop virus, increased viral burden (more HA) correlates with IFN induction.
A similar trend also holds for the virus with an amino-acid substitution in NS1 (NS1-A122V). 
But for wildtype and the two PB1 mutants, there is no clear trend for higher viral burden to correlate with IFN induction.
\FIG{mutantcomparison}A summarizes the data in this figure by simply showing the ratio of the percent IFN+ of the highest HA-expression quartile to the lowest HA-expression quartile for each sample and replicate.}
{\includegraphics[width=0.9\textwidth]{figures/MutantComparison/p_ifn_vs_ha.pdf}}
\label{figsupp:ha_vs_ifn}

\figsupp[Similar to \FIG{mutantcomparison}B, but uses cycloheximide rather than ribavirin.]
{Preventing viral secondary transcription using cycloheximide also decreases IFN induction by the NS1stop virus relative to the PB1 deletion virus, corroborating the result in \FIG{mutantcomparison}B with a different inhibitor.
However, the results are more complex here, since cycloheximide alone increases IFN production~\citep{raj1981analysis, maroteaux1983cycloheximide, killip2014activation}.
Therefore, we see an increase in IFN levels upon cycloheximide treatment for uninfected cells and cells infected with the PB1 deletion virus, but a decrease for cells infected with the NS1stop virus.
This plot shows qPCR results for the same samples as in \FIGSUPP[validation]{cycloheximide}, but with the qPCR for expression of IFNB1 transcripts relative to the housekeeping gene L32.
As described in \FIGSUPP[validation]{cycloheximide}, A549 cells were infected at an MOI of 0.4, such that HA expression was equivalent between viral variants when secondary transcription was blocked, and some of the samples treated with 50 $\mu$g/ml of cycloheximide~\citep[a concentration sufficient to block secondary transcription;][]{killip2014activation}.
After 8 hours, transcript levels were quantified by qPCR.
Points indicate four replicates, and lines indicate the mean.
The dashed line shows the limit of detection, and points below this limit where set to that value.
}
{\includegraphics[width=0.6\textwidth]{figures/MutantComparison/cyclo_qPCR_plot.pdf}}
\label{figsupp:cycloheximide}
